---
title: "CDF with Proteomics"
author: "Bing Shui"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Library set-up
```{r message=FALSE, warning=FALSE}
library(DESeq2)
library(data.table)
library(gdata)
library(ggplot2)
library(RColorBrewer)
library(tidyverse)
library(plotly)
library(AnnotationDbi)
library(org.Mm.eg.db)
```

# Purpose
 - To examine if up/down-regulated proteins from KRas colonic tumor proteomics preferentially have up/down-regulated ATAC-Seq peaks
 - To examine if up/down-regulated ATAC-Seq peaks preferrentially have up/down-regulated proteins from KRas colonic tumor proteomics

# Set up function
```{r}
plotCDF.ggplot <- function(gene.counts, gene.sets, gene.set.labels,
                           col = "", linetype = "", xlim = c( -1.0, 1.3 ),
                           legend.size = 22, axistitle.size = 22, title = "Fold change log2 (Dicer KO/WT)",
                           legend.pos = c(0.7, 0.18), geom = "step", ploc = -0.9) {
  require(ggplot2)
  df.log2fc <- gene.counts[,c("gene", "log2FC")]
  #rownames(df.log2fc) <- df.log2fc$gene
  if (length(gene.sets) != length(gene.set.labels)){
    return("Length of gene sets doesn't match labels")
  }
  target.expr <- df.log2fc[df.log2fc$gene %in% gene.sets[[1]],]
  for (i in 2:length(gene.sets)){
    target.expr <- rbind(target.expr, df.log2fc[df.log2fc$gene %in% gene.sets[[i]],])
  }

  gene.set.counts <- c()
  for (j in 1:length(gene.sets)){
    gene.set.counts <- c(gene.set.counts, sum(df.log2fc$gene %in% gene.sets[[j]]))
  }
  
  target.expr$Category <- rep(gene.set.labels, gene.set.counts)
  target.expr$Category <- factor(target.expr$Category, levels = gene.set.labels)
  
  log2FC.values <- lapply(gene.sets, function(gene.set) {
    gene.counts[gene.counts$gene %in% gene.set,]$log2FC
  })

  ks.pvals <- lapply(log2FC.values,
                     function(log2FCs) {
                       ks.test(log2FCs, log2FC.values[[1]])$p.value
                     })
  
  p <- ggplot( target.expr, aes( x = log2FC, colour = Category ) ) +
  stat_ecdf( geom = geom, aes( colour = Category, linetype = Category ), lwd = 1 ) +
  scale_color_manual( values = col, labels = sprintf( "%s (%d)", gene.set.labels, gene.set.counts ) ) +
  scale_linetype_manual( values = linetype, labels = sprintf( "%s (%d)", gene.set.labels, gene.set.counts ) ) +
  # xlim() will remove data points; Be careful in the future
  coord_cartesian( xlim = xlim ) + xlab(title) + ylab('CDF') +
  theme_classic() + theme( legend.background = element_rect(fill = NA), 
                           legend.title = element_blank(), 
                           legend.position = legend.pos,
                           legend.text = element_text(size=legend.size),
                           legend.key.size = unit(1.5, 'lines'),
                           axis.title.x = element_text(size=axistitle.size, margin = margin(t = 10)),
                           axis.title.y = element_text(size=axistitle.size, margin = margin(r = 10)),
                           axis.text=element_text(size=20),
                           axis.line = element_line(size = 1),  #axis label size
                           axis.ticks.length = unit(0.3, "cm")) #increase tick length
  
  for (k in 2:length(gene.sets)){
    p <- p + annotate(geom = "text", x = ploc, y = 1-0.08*(k-1), hjust = 0, 
                      label = sprintf("p = %.0e", ks.pvals[k]), 
                      colour = col[k], size = 8)
  }
  print(p)
}
```
# Load datasets
```{r message=FALSE, warning=FALSE}
# Proteomics
protein_DGE <- read_csv("../../../../Proteomics data/colon tumor-enema model/crcMS_diff.csv")

annotations <- AnnotationDbi::select(org.Mm.eg.db,
                                           keys = protein_DGE$`Protein Id`,
                                           columns = c("ENSEMBL"),
                                           keytype = "UNIPROT")

# Determine the indices for the non-duplicated genes
non_duplicates <- which(duplicated(annotations$UNIPROT) == FALSE)

# Return only the non-duplicated genes using indices
annotations <- annotations[non_duplicates, ]

# Check number of NAs returned
is.na(annotations$UNIPROT) %>%
  which() %>%
  length()

protein_DGE <- left_join(protein_DGE, annotations, by = c("Protein Id"="UNIPROT"))

colnames(protein_DGE)[8] <- "gene"
colnames(protein_DGE)[7] <- "log2FC"

# ATAC-Seq
atac_res <- read_csv("Result/ATAC_crc_Kras-WT_merged.csv")
colnames(atac_res)[24] <- "gene"
colnames(atac_res)[8] <- "log2FC"
```

# Sig ATAC-Seq peaks in Proteomics
## Up-regulated ATAC-Seq peaks
I expect general up-regulation of their target RNA
```{r}
cols <- c(brewer.pal(name = "Set2", n = 4))

atac.up <- atac_res %>% filter(padj < 0.05 & log2FC > 0) %>% dplyr::select(gene) %>% unlist()
atac.up <- atac.up[!is.na(atac.up)]
plotCDF.ggplot(protein_DGE,
               list(protein_DGE$gene, atac.up),
               c("All genes", "Genes with ATAC-UP"),
               col = c("grey15", cols[1]),
               linetype = c(1, 1),
               title = "Protein_LFC(K-Ras_G12D/WT)",
               xlim = c(-1,2),
               legend.size = 15)

pdf("PDF_Figure/CDF_Gene with ATAC-UP in Proteomics.pdf",
    width = 10,
    height = 8)
plotCDF.ggplot(protein_DGE,
               list(protein_DGE$gene, atac.up),
               c("All genes", "Genes with ATAC-UP"),
               col = c("grey15", cols[1]),
               linetype = c(1, 1),
               title = "Protein_LFC(K-Ras_G12D/WT)",
               xlim = c(-1,2),
               legend.size = 15)
dev.off()
```

## Down-regulated ATAC-Seq peaks
I expect general down-regulation of their target RNA
```{r}
atac.down <- atac_res %>% filter(padj < 0.05 & log2FC < 0) %>% dplyr::select(gene) %>% unlist()
atac.down <- atac.up[!is.na(atac.down)]

plotCDF.ggplot(protein_DGE,
               list(protein_DGE$gene, atac.down),
               c("All genes", "Genes with ATAC-DOWN"),
               col = c("grey15", cols[2]),
               linetype = c(1, 1),
               title = "Protein_LFC(K-Ras_G12D/WT)",
               xlim = c(-1,2),
               legend.size = 15)

pdf("PDF_Figure/CDF_Gene with ATAC-DOWN in Proteomics.pdf",
    width = 10,
    height = 8)
plotCDF.ggplot(protein_DGE,
               list(protein_DGE$gene, atac.down),
               c("All genes", "Genes with ATAC-DOWN"),
               col = c("grey15", cols[2]),
               linetype = c(1, 1),
               title = "Protein_LFC(K-Ras_G12D/WT)",
               xlim = c(-1,2),
               legend.size = 15)
dev.off()
```

# DEG from Proteomics in ATAC-Seq peaks
## Up-regulated proteins in Proteomics 
I expect general Up-regulation of these genes' ATAC-Seq peaks
```{r}
protein.up <- protein_DGE %>% filter(p_values < 0.05 & q_values < 0.1 & log2FC > 0) %>% dplyr::select(gene) %>% unlist()
protein.up <- protein.up[!is.na(protein.up)]

plotCDF.ggplot(atac_res,
               list(atac_res$gene, protein.up),
               c("All genes", "Genes with Protein-UP"),
               col = c("grey15", cols[3]),
               linetype = c(1, 1),
               title = "ATAC_LFC(K-Ras_G12D/WT)",
               xlim = c(-0.5,0.5),
               legend.size = 15,
               geom = "line",
               ploc = -0.5)

pdf("PDF_Figure/CDF_Gene with Protein-UP in ATACSeq.pdf",
    width = 10,
    height = 8)
plotCDF.ggplot(atac_res,
               list(atac_res$gene, protein.up),
               c("All genes", "Genes with Protein-UP"),
               col = c("grey15", cols[3]),
               linetype = c(1, 1),
               title = "ATAC_LFC(K-Ras_G12D/WT)",
               xlim = c(-0.5,0.5),
               legend.size = 15,
               geom = "line",
               ploc = -0.5)
dev.off()
```

## Down-regulated proteins in Proteomics 
I expect general Down-regulation of these genes' ATAC-Seq peaks
```{r}
protein.down <- protein_DGE %>% filter(p_values < 0.05 & q_values < 0.1 & log2FC < 0) %>% dplyr::select(gene) %>% unlist()
protein.down <- protein.down[!is.na(protein.down)]

plotCDF.ggplot(atac_res,
               list(atac_res$gene, protein.down),
               c("All genes", "Genes with Protein-DOWN"),
               col = c("grey15", cols[4]),
               linetype = c(1, 1),
               title = "ATAC_LFC(K-Ras_G12D/WT)",
               xlim = c(-0.5,0.5),
               legend.size = 15,
               geom = "line",
               ploc = -0.5)

pdf("PDF_Figure/CDF_Gene with Protein-DOWN in ATACSeq.pdf",
    width = 10,
    height = 8)
plotCDF.ggplot(atac_res,
               list(atac_res$gene, protein.down),
               c("All genes", "Genes with Protein-DOWN"),
               col = c("grey15", cols[4]),
               linetype = c(1, 1),
               title = "ATAC_LFC(K-Ras_G12D/WT)",
               xlim = c(-0.5,0.5),
               legend.size = 15,
               geom = "line",
               ploc = -0.5)
dev.off()
```

# SessionInfo
```{r}
sessionInfo()
```